Lawvere-Tierney Sheafification\\ in Homotopy Type Theory

### 2015-01-16 paper/sheaf_lics.tex
pass on section 2

\subsection{The epi-mono Factorization System}
\label{sec:epi-mono-fact}

Now, let a map $f:A\to B$ be called {\em (-1)-truncated} if
every fibers of $f$ is an $\HProp$, and {\em
  (-1)-connected} if every fibers $f$ is merely
inhabited, \ie such that $\|\fib f y\|$ for all $y$.

\begin{defi}
  Let $f:A\to B$ be a function. The {\em image} of $f$ is
  $$\im(f) = \sum_{y:B} \|\fib f y\|.$$
\end{defi}

It is then fairly obvious that the canonical function $\hat f : A \to
\im(f)$ is (-1)-connected, thus the following proposition arise
(\cite{hottbook}, Lemma 7.6.4)~:
\begin{prop}
  An map $f:A\to B$ factors as a (-1)-connected function followed by a
  (-1)-truncated function.
\end{prop}

\nt{define the image here, with (-1)-truncated and (-1)-connected
  factorization system}

@@

Then $P$ is the {\em colimit} of $D$ if for any other $X:\Type$, the
map
$$\lambda f:P \to X,~ \left( \lambda i,~f \circ q_i~;~ \lambda i\, j\,
  \phi,~ \postcompose_f (p(i, j, g))) \right),$$
where $\postcompose_f : \phi = \psi \to \phi \circ f = \psi \circ f$,
is an equivalence.
\end{defi}

@@

% We use the following definition of truncated left exact modalities:
\begin{defi}
  \label{sec:defin-basic-prop-1}
  Let $n\geq -1$ be a truncation index. A left exact modality at level
  $n$ is the data of
  \kq{Do you prefer $n$ over $p$ ?}
  \begin{enumerate}[(i)]
  \item A predicate $P:\Type_p \to \HProp$
  \item For every $n$-truncated type $A$, a $n$-truncated type
    $\modal A$ such that $P(\modal A)$
  \item For every $n$-truncated type $A$, a map $\eta_A:A \to
    \modal A$
  \end{enumerate}
  such that
  \begin{enumerate}[(i)]
    \setcounter{enumi}{3}
  \item For every $n$-truncated types $A$ and $B$, if $P(B)$ then
    $$\left\{
      \begin{array}{rcl}
        (\modal A \to B) & \to & (A \to B) \\
        f & \mapsto & f \circ \eta_A
      \end{array} \right.$$
    is an equivalence.
  \item for any $A:\Type_p$ and $B:A \to \Type_p$ such that $P(A)$
    and $\prod_{x:A} P(B x)$, then $P\left( \sum_{x:A} B(x)\right)$
  \item for any $A:\Type_p$ and $x,y:A$, if $\modal A$ is
    contractible, then $\modal (x=y)$ is contractible.
  \end{enumerate}
  Conditions (i) to (iv) define a {\em reflective subuniverse}, (i) to
  (v) a {\em modality}.
\end{defi}

The type of all modal $n$-types, \ie types for which $P$ holds, is
defined as %$\Type^\modal_p$, so 
$$
\Type^\modal_p \ \defeq \sum_{T : \Type_n} P (T).
$$
%
Since basic type formers (dependent products, products, sigma types)
preserves truncation levels, we can leverage all theorems on modalites
presented in~\cite[Chapter 7.7]{hottbook}.

We now summarize the main properties that are used in our
formalization.

\begin{prop}
\vspace{1em}

\vspace{1em}

@@

\item $\one$ seen as a $\Type_p$ is modal, ie $\modal \one =\one$.
\item (\cite{hottbook}, Lemma 7.7.2)
  Let $A:\Type$ and $B:A \to \Type_p^\modal$. Then $P\left(\prod_{a:A} (B\,
  a)\right)$. 
\item (\cite{hottbook}, Corollary 7.7.2) For any $A,B:\Type_p$,
  $\modal (A\times B) = (\modal A) \times (\modal B)$.
\item Let $A:\Type_p\modal$. Then for any $x,y:X$, $P(x=y)$.
\item (\cite{hottbook}, Theorem 7.7.4) 
  {\it (iv)} can be extended to dependent product. Precisely~:

  Let $A:\Type_p$ and $B:\modal A \to \Type_p^\modal$. For any map
  $g:\prod_{a:A} B(\eta_A(a))$ there exists a map $f:\prod_{z:\modal
    A} (B\, z)$ such that for all $a:A$, $f(\eta_A(a)) = g(a)$.
\end{itemize}
\end{prop}

@@

Moreover the following diagram commute
$$\xymatrix{
  \sum_{x:X} (f x = y) \ar[r]^\eta \ar[d]_\gamma & \modal \left(\sum_{x:X}  (f x = y)\right) \ar@{=}[dl]\\
  \sum_{x:\modal X} (\modal f x = \eta_Y y) &
}$$
where $\pi_1 \circ \gamma = \eta_X \circ \pi_1$, and $\pi_2 \circ
\gamma$ is given by the modalisation of paths.
\end{prop}
\begin{proof}
\nt{ça a du sens de mettre une preuve, non ? c'est non trivial et
  nouveau par rapport aux travaux sur les modalités. Ou alors c'est
  la preuve de Shulman qui tu n'as pas faite. En tout cas, il faut
  clarifier.}

The central result used here is the following
\begin{lem}
  Let $X:\Type_p$, $Y:\Type_p\modal$ and $f:X\to Y$. If for all $y:Y$,
  $\modal (\fib f y)$ is contractible, then $\modal X = Y$.
\end{lem}

Here, $\sum_{x:\modal X} (\modal f x = \eta_Y y)$ is modal, thus we
can define $$\phi:\modal \left(\sum_{x:X}  (f x = y)\right) \to
\sum_{x:\modal X} (\modal f x = \eta_Y y).$$
We just need to check that the image of its fibers by $\modal$ are
contractible.

From definition of left exactness, one can deduce the following~:
\begin{lem}
  Let $A,B:\Type_p$. Let $f:A\to B$. If $\modal A$ and $\modal B$ are
  contractible, then so is $\fib f b$ for any $b:B$.
\end{lem}
We can rewrite $\fib \phi {x;p}$ in terms of the fiber of a fonction
$\fib {\eta_X} a \to \fib {\eta_Y} b$.

>>>>

\subsection{The Epi-Mono Factorization System}
\label{sec:epi-mono-fact}

We now recall the presence of a factorization system in homotopy type
theory, constituted by monomorphism and epimorphism.

A map $f:A\to B$ is a \emph{monomorphism} when it is (-1)-truncated
(see definition in Section~\ref{ssec:hott}), and an \emph{epimorphism}
when it is (-1)-connected, that is when every fiber of $f$ is merely
inhabited (\ie $\|\fib f y\|$ holds for all $y$).

\begin{defi}
  Let $f:A\to B$ be a function. The {\em image} of $f$ is defined as 
  $$\im(f) \defeq \sum_{y:B} \|\fib f y\|.$$
\end{defi}

It is then fairly obvious that the canonical function $\hat f : A \to
\im(f)$ is (-1)-connected, thus the following proposition holds
(\cite{hottbook}, Lemma 7.6.4)~:
\begin{prop}
  A map $f:A\to B$ factors through $\im(f)$ as a (-1)-connected
  function followed by a (-1)-truncated function.
\end{prop}

Note that there are other factorization systems constituted by
$n$-truncated morphism and $n$-connected morphisms, for every
truncated level $n$, but they do not appear to be useful in the
definition of sheafification.

@@

Then $P$ is the {\em colimit} of $D$ if for any other $X:\Type$, the
map
$$\lambda f:P \to X, \left( \lambda i,~f \circ q_i~;~ \lambda i\, j\,
  \phi,~ \postcompose_f (p(i, j, g)) \right)$$
where $\postcompose_f : \phi = \psi \to \phi \circ f = \psi \circ f$,
is an equivalence.
\end{defi}

@@

% We use the following definition of truncated left exact modalities:
\begin{defi}
  \label{sec:defin-basic-prop-1}
  Let $n\geq -1$ be a truncation index. A left exact modality at level
  $n$ is the data of
  \begin{enumerate}[(i)]
  \item A predicate $P:\Type_n \to \HProp$
  \item For every $n$-truncated type $A$, a $n$-truncated type
    $\modal A$ such that $P(\modal A)$
  \item For every $n$-truncated type $A$, a map $\eta_A:A \to
    \modal A$
  \end{enumerate}
  such that
  \begin{enumerate}[(i)]
    \setcounter{enumi}{3}
  \item For every $n$-truncated types $A$ and $B$, if $P(B)$ then
    $$\left\{
      \begin{array}{rcl}
        (\modal A \to B) & \to & (A \to B) \\
        f & \mapsto & f \circ \eta_A
      \end{array} \right.$$
    is an equivalence.
  \item for any $A:\Type_n$ and $B:A \to \Type_n$ such that $P(A)$
    and $\prod_{x:A} P(B x)$, then $P\left( \sum_{x:A} B(x)\right)$
  \item for any $A:\Type_n$ and $x,y:A$, if $\modal A$ is
    contractible, then $\modal (x=y)$ is contractible.
  \end{enumerate}
  Conditions (i) to (iv) define a {\em reflective subuniverse}, (i) to
  (v) a {\em modality}.
\end{defi}

The type of all modal $n$-types, \ie types for which $P$ holds, is
defined as %$\Type^\modal_p$, so 
$$
\Type^\modal_n \ \defeq \sum_{T : \Type_n} P (T).
$$
%
Since basic type formers (dependent products, products, sigma types)
preserves truncation levels, we can leverage all theorems on modalites
presented in~\cite[Chapter 7.7]{hottbook}.
%
We now summarize the main properties that are used in our
formalization. We do not recall the proofs as they are not specific to
our presentation (also the setting using truncated types provides
minor changes), but they can be found in our Coq form

\begin{prop}

@@

\item $\one$ seen as a $\Type_p$ is modal, ie $\modal \one =\one$.
\item 
  Let $A:\Type$ and $B:A \to \Type_n^\modal$. Then $P\left(\prod_{a:A} (B\,
  a)\right)$ (from \cite[Lemma 7.7.2]{hottbook}). 
\item For any $A,B:\Type_p$,
  $\modal (A\times B) = (\modal A) \times (\modal B)$ \\
(from \cite[Corollary 7.7.2]{hottbook}).
\item Let $A:\Type_n \modal$. Then for any $x,y:X$, $P(x=y)$.
\item 
  {\it (iv)} can be extended to dependent product, that is for any $A:\Type_p$, $B:\modal A \to
  \Type_p^\modal$ and
  $g:\prod_{a:A} B(\eta_A(a))$, there exists a map $f:\prod_{z:\modal
    A} (B\, z)$ such that for all $a:A$, $f(\eta_A(a)) = g(a)$ \\
 (from \cite[Theorem
 7.7.4]{hottbook}).
\end{itemize}
\end{prop}

@@

Moreover the following diagram commute
$$\xymatrix{
  \sum_{x:X} (f x = y) \ar[r]^\eta \ar[d]_\gamma & \modal \left(\sum_{x:X}  (f x = y)\right) \ar@{=}[dl]\\
  \sum_{x:\modal X} (\modal f x = \eta_Y y) &
}$$
where $\pi_1 \circ \gamma = \eta_X \circ \pi_1$, and $\pi_2 \circ
\gamma$ is given by the modalisation of paths.
\end{prop}
\begin{proof}
% \nt{ça a du sens de mettre une preuve, non ? c'est non trivial et
%   nouveau par rapport aux travaux sur les modalités. Ou alors c'est
%   la preuve de Shulman qui tu n'as pas faite. En tout cas, il faut
%   clarifier.}

The central result used here is the following
\begin{lem}
  Let $X:\Type_p$, $Y:\Type_p\modal$ and $f:X\to Y$. If for all $y:Y$,
  $\modal (\fib f y)$ is contractible, then $\modal X = Y$.
\end{lem}
%
Here, $\sum_{x:\modal X} (\modal f x = \eta_Y y)$ is modal, thus we
can define $$\phi:\modal \left(\sum_{x:X}  (f x = y)\right) \to
\sum_{x:\modal X} (\modal f x = \eta_Y y).$$
We just need to check that the image of its fibers by $\modal$ are
contractible.
%
From definition of left exactness, one can deduce the following~:
\begin{lem}
  Let $A,B:\Type_p$. Let $f:A\to B$. If $\modal A$ and $\modal B$ are
  contractible, then so is $\fib f b$ for any $b:B$.
\end{lem}
We can rewrite $\fib \phi {x;p}$ in terms of the fiber of a fonction
$\fib {\eta_X} a \to \fib {\eta_Y} b$.
